\Large{Adversarial Variational Optimization of Non-Differentiable Simulators}

### 2017-07-14 tex/paper.tex
save

The variational objectives \ref{eqn:vo-ud}-\ref{eqn:vo-ug}
have the effect of replacing the modeled data distribution of Eqn.~\ref{eqn:p_theta} with
the marginal distribution parametrized by $\bfpsi$
\begin{equation}
    \qxpsi = \int  p(\mathbf{x}|\bftheta) q(\bftheta|\bfpsi) d\bftheta.
\end{equation}
We can think of $q(\bfx|\bfpsi)$ as a \textit{variational program} as described
in \citep{2016arXiv161009033R}, though more complicated than a simple
reparameterization of normally distributed noise $\bfz$ through a differentiable
function $g(\bfz, \bfpsi)$. In our case, the variational program is a
marginalized, non-differentiable  simulator.  Its density is  intractable;
nevertheless, it can generate samples for $\bfx$ and is differentiable with
respect to $\bfpsi$.
Operationally, we sample from this marginal model via
\begin{equation}\label{eqn:p_psi}
    \mathbf{x} \sim \qxpsi \equiv \bftheta \sim q(\bftheta|\bfpsi), \bfz \sim p(\bfz|\bftheta), \mathbf{x} = g(\bfz; \bftheta).
\end{equation}

@@

Typically in variational inference one starts with a tractable prior
$p(\bftheta)$ and a tractable likelihood $p(\bfx |\bftheta )$, thus fixing the
posterior $p(\bftheta | \bfx )$.   After optimizing $q(\bftheta | \bfpsi)$ to
approximate the posterior, one would criticize the model through the the
posterior predictive distribution as advocated by Box~\citep{box1980sampling}.
Effectively, we are optimizing variational posterior $q(\bftheta | \bfpsi)$ by
directly minimizing the  divergence of the corresponding posterior predictive
distribution $\qxpsi$ and the data distribution $p_r(\bfx)$ without evaluating
the intractable likelihood $p(\bfx | \bftheta )$ or specifying a prior.
Consequently, the proposed algorithm is a form of variational inference through
model criticism~\citep{box1980sampling}, which we denote AVO-EB. While the
likelihood is intractable, it is fixed, thus one can view the optimization of
$\bfpsi$ through the lens of empirical Bayes, though the prior family
$p(\bftheta | \bfpsi)$ is not explicitly specified.

>>>>

The variational objectives \ref{eqn:vo-ud}-\ref{eqn:vo-ug}
have the effect of replacing the modeled data distribution of Eqn.~\ref{eqn:p_theta} with
the parameterized marginal distribution of the generated data
\begin{equation}
    \qxpsi = \int  p(\mathbf{x}|\bftheta) q(\bftheta|\bfpsi) d\bftheta.
\end{equation}
We can think of $q(\bfx|\bfpsi)$ as a \textit{variational program} as described
in \citep{2016arXiv161009033R}, though more complicated than a simple
reparameterization of normally distributed noise $\bfz$ through a differentiable
function. In our case, the variational program is a
marginalized, non-differentiable  simulator.  Its density is  intractable;
nevertheless, it can generate samples for $\bfx$ and is differentiable with
respect to $\bfpsi$.
Operationally, we sample from this marginal model via
\begin{equation}\label{eqn:p_psi}
    \mathbf{x} \sim \qxpsi \equiv \bftheta \sim q(\bftheta|\bfpsi), \bfz \sim p(\bfz|\bftheta), \mathbf{x} = g(\bfz; \bftheta).
\end{equation}

@@

Typically in variational inference one starts with a tractable prior
$p(\bftheta)$ and a tractable likelihood $p(\bfx |\bftheta )$, thus fixing the
posterior $p(\bftheta | \bfx )$.   After optimizing $q(\bftheta | \bfpsi)$ to
approximate the posterior, one would criticize the model through the the
posterior predictive distribution as advocated by Box~\citep{box1980sampling}.
Effectively, we are optimizing the variational posterior $q(\bftheta | \bfpsi)$ by
directly minimizing the  divergence of the corresponding posterior predictive
distribution $\qxpsi$ and the data distribution $p_r(\bfx)$ without evaluating
the intractable likelihood $p(\bfx | \bftheta )$ or specifying a prior.
Consequently, the proposed algorithm is a form of variational inference through
model criticism~\citep{box1980sampling}, which we denote AVO-EB. While the
likelihood is intractable, it is fixed, thus one can view the optimization of
$\bfpsi$ through the lens of empirical Bayes, though the prior family
$p(\bftheta | \bfpsi)$ is not explicitly specified.
